%>% -> piping operator
Everything that comes before the pipe is evaluated and passed as first argument into the function that comes after the pipe

```{r}
install.packages('readr')
install.packages('dplyr')
install.packages('rlang')
install.packages('pillar')

library(readr)
library(dplyr)
system.file(package = 'dplyr')

bill <- read_csv("charts.csv")

bill %>% tail(10)
10 %>% sample_n(bill,.)

```

dplyr can be grouped into categories based on 
    1. column operation - select( ), mutate( )
    2. row operation - filter( ), distinct( ), arrange( )
    3. group operation - group_by( ), summarise( ), count( )
     
select( ) -> select columns by name or helper functions


# select( )

```{r}

bill = bill [-10,]
bill

bill %>% 
  select(.,date,rank,song,artist, 'weeks-on-board' )

bill %>% 
  select(.,date:artist, weeks_popular = 'weeks-on-board')
bill

bill %>% 
  select(., -`last-week`, -`peak-rank`)

```

helper functions
```{r}
select(bill, contains("art", ignore.case = TRUE))
# select columns contains the .

select(bill, ends_with("ard"))
# select columns ends with .

select(bill, everything())
# select every columns

select(bill, matches("."))
# select columns that matches regex

select(bill, num_range("x", 1:5))
# select columns named x1,x2,x3,x4,x5

select(bill, one_of(c('artist','not')))
# select columns whose names are in group of names

select(bill, starts_with("art"))
# select columns that start with this character

```


mutate() -> compute ot append one or more new columns

```{r}
bill


bill %>% 
  mutate(iscollab = grepl("Featuring", artist)) %>% 
  select(artist,iscollab, week_popular=`weeks-on-board`)
```

filter() ->Extract rows that meet logical criteria

```{r}
bill %>% 
  select(song,rank,artist, weeks_popular = `weeks-on-board`) %>% 
  filter(.,weeks_popular>20, artist == 'Adele' | artist == 'Coldplay')

bill %>% 
  select(song,rank,artist,week_popular = `weeks-on-board`) %>% 
  filter(., week_popular>20, grepl("^lil",ignore.case = T, artist))


```

distinct() -> remove duplicate rows

```{r}
bill %>% 
  select(.,artist,song,rank) %>% 
  filter(artist == 'Coldplay') %>% 
  distinct(song)
  
  #this returns as dataframe

bill %>% 
  select(.,artist,song,rank) %>% 
  filter(artist == 'Coldplay') %>% 
  distinct(song) %>% 
  .$song # fetch the df object and take the colummn song


```

group_by() -> group data into rows with the same value
summarise() -> summarise data into single row of values

```{r}

bill %>% 
  select(., artist, song, rank, weeks_popular='weeks-on-board') %>% 
  filter(artist == 'Coldplay')  %>% 
  group_by(song) %>% 
  summarise(total_weeks_popular = max(weeks_popular)) 
  
```

arrange() -> order rows by values of a column [low to high]


```{r}
bill %>% 
  select(artist, rank, song, pop = `weeks-on-board`) %>% 
  filter(artist == 'Coldplay') %>% 
  group_by(song) %>% 
  summarise(total_weeks_pop = max(pop)) %>% 
  arrange(-total_weeks_pop,song) %>% 
  head(10)
```

count() -> count number of rows with each unique value of the variable

```{r}
bill %>% 
  select(artist, rank, song, week_pop = `weeks-on-board`) %>% 
  count(artist) %>% 
  rename(times = n) %>% 
  arrange(-times)

```











