SAI

Data cleaning (lubridate & stringr)
Reshaping Data (tidyr)
Joining / Merging Datasets


```{r}
install.packages('lubridate')
install.packages('stringr')

library(readxl)
library(dplyr)
library(lubridate)
library(stringr)



bill = read.csv("charts.csv")
bill 

musicdf = bill %>% 
  select(date:artist, weeks_popular = 'weeks.on.board')
musicdf
```

lubridate package -> deal with the date and time values -> clean and parse date and time
stringr package ->string extraction and matching

```{r}
musicdf %>% 
  mutate(date = ymd(date))

```

stringr package str_detect, str_match and other useful string functions

str_detect(string, pattern) -> Detect the presence of a pattern match in a string
str_match(string, pattern) -> Return the first pattern  match found in each string, as a matrix with a column for each() group in pattern.
str_match_all(sentence, pattern)



```{r}

musicdf <- musicdf %>% 
  mutate(date=ydm(date),
  primary_artist = ifelse(str_detect(artist, 'Featuring'),    #ifelse(cond, T,F)
                          str_match(artist, '(.*)\\sFeaturing')[,2],
                          artist),
  featured_artist = str_match(artist, 'Featuring\\s(.*)')[,2]) %>% 
  select(date,artist, primary_artist,featured_artist)

musicdf



# [,2] using because 
# full match is going to be the 1st thing that's returned
```















































